---
title: "Australia CO2 emission"
author: "Aphiaut Imuan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(GGally)
library(mgcv)
library(broom)
library(visreg)
library(patchwork)
```

```{r, warning=FALSE, echo=FALSE, include=FALSE}
co2 <- read_csv("data/co-emissions-per-capita.csv") %>%
  filter(Entity == "Australia",
         Year >= 1996) %>%
  rename("year" = "Year",
         "co2_capita" = "Annual CO₂ emissions (per capita)") %>%
  select(year, co2_capita)


pop <- read_csv("data/API_SP.POP.TOTL_DS2_en_csv_v2_5551506.csv", skip = 4) %>%
  pivot_longer(cols = 5:67, names_to = "year", values_to = "total_pop") %>%
  rename("country" = "Country Name") %>%
  filter(country == "Australia",
         year >= 1996) %>%
  select(c(year, total_pop)) %>%
  na.omit() %>%
  mutate(year = as.numeric(year))


migrant <- read_csv("data/API_SM.POP.NETM_DS2_en_csv_v2_5551569.csv", skip = 4) %>%
  pivot_longer(cols = 5:67, names_to = "year", values_to = "net_migrant") %>%
  rename("country" = "Country Name") %>%
  filter(country == "Australia",
         year >= 1996,
         year < 2022) %>%
  select(c(year, net_migrant)) %>%
  na.omit() %>%
  mutate(year = as.numeric(year))


gdp <- read_csv("data/API_NY.GDP.PCAP.KD_DS2_en_csv_v2_5551580.csv", skip = 4) %>%
  pivot_longer(cols = 5:67, names_to = "year", values_to = "gdp_capita") %>%
  rename("country" = "Country Name") %>%
  filter(country == "Australia",
         year >= 1996) %>%
  select(c(year, gdp_capita)) %>%
  na.omit() %>%
  mutate(year = as.numeric(year))


myCols <- as.character(read_excel("data/Australian Energy Statistics 2021 Table B_0.xlsx", sheet = "AUS", skip = 4, n_max = 1, col_names = FALSE))
myCols <- c("year", myCols)

energy <- read_excel("data/Australian Energy Statistics 2021 Table B_0.xlsx", sheet = "AUS", skip = 6, col_names = myCols) %>% 
  na.omit() %>%
  filter(as.numeric(sub("-.*", "", year)) %in% 1996:2021) %>%
  mutate(year = as.numeric(sub("-.*", "", year))) %>%
  select(c("year", "Energy consumption")) %>%
  rename("energy_consumption" = "Energy consumption")

co2_new<- co2 %>%
  right_join(pop, by = "year") %>%
  right_join(migrant, by = "year") %>%
  right_join(gdp, by = "year") %>%
  right_join(energy, by = "year")

```

# Data Description

CO2 emission from [Our World in Data](https://ourworldindata.org/co2/country/australia) This data focus on 25 years 1996-2021

Energy data from  [Department of Climate Change, Energy, the Environment and Water](https://www.energy.gov.au/publications/australian-energy-update-2021)

Population data from  [World Bank database](https://data.worldbank.org/indicator/SP.POP.TOTL?locations=AU)

Net migration data from  [World Bank database](https://data.worldbank.org/indicator/SM.POP.NETM?locations=AU)

GDP per capita constant 2015$ US [World Bank database](https://data.worldbank.org/indicator/NY.GDP.MKTP.KD?locations=AU)

# Theory for this report
IPAT theory or The Environmental Impacts (I) on Population (P), Affluence (A), and Technology (T)
EKC theory or the environmental Kuznets curve created by Simon Smith Kuznets


```{r, warning=FALSE, echo=FALSE}
ggpairs(co2_new)
```

From the figure illustrates correlations between CO2 and predictors are not linear relationship, therefore all predictors need to transform. In this transformation based on the Stochastic Regression on Population, Affluence and Technology (STIRPAT) framework because the IPAT model is theory that not useful for statistic. However, this framework has a drawback that the correlation between response variable and predictors are linear relationship. 

```{r, echo=FALSE}
co2_trans <- co2_new %>%
  mutate(total_pop = log(total_pop),
         net_migrant = log(net_migrant),
         gdp_capita = log(gdp_capita),
         energy_consumption = log(energy_consumption),
         co2_capita = log(co2_capita))
```

```{r, echo=FALSE, warning=FALSE}
co2_trans %>%
  dplyr::select(-year) %>%
  GGally::ggpairs(aes(alpha=.1))
```


```{r, echo=FALSE, warning=FALSE}
lmod_1 <- lm(co2_capita ~ total_pop + net_migrant + gdp_capita + energy_consumption + total_pop:net_migrant + total_pop:gdp_capita + total_pop:energy_consumption, data = co2_trans)

summary(lmod_1)
```
```{r, echo=FALSE, warning=FALSE}
broom::augment(lmod_1) %>%
  ggplot() +
  geom_qq(aes(sample=.resid))
```


All p-value of predictors are not significant. It means they do not have enough evidence to conclude that predictors affect to CO2 emission per capita in Australia. Moreover, this QQ-plot shows outliers from both tails, and a close alignment of points is not linear line.

```{r, echo=FALSE, warning=FALSE}
co2_pop <- co2_trans %>%
  ggplot(aes(x = total_pop, y = co2_capita)) +
  geom_smooth() + 
  geom_point() +
  labs(y = "CO2 per capita", x = "Total population", title = "Relationship between CO2 per capita and Total population") +
  theme(plot.title = element_text(size = 7.5))

co2_migrant <- co2_trans %>%
  ggplot(aes(x = net_migrant, y = co2_capita)) +
  geom_smooth() + 
  geom_point() +
  labs(y = "CO2 per capita", x = "Net migrants", title = "Relationship between CO2 per capita and Net migrants") +
  theme(plot.title = element_text(size = 7.5))

co2_gdp <- co2_trans %>%
  ggplot(aes(x = gdp_capita, y = co2_capita)) +
  geom_smooth() + 
  geom_point() +
  labs(y = "CO2 per capita", x = "GDP per capita", title = "Relationship between CO2 per capita and GDP per capita") +
  theme(plot.title = element_text(size = 7.5))

co2_ene <- co2_trans %>%
  ggplot(aes(x = energy_consumption, y = co2_capita)) +
  geom_smooth() + 
  geom_point() +
  labs(y = "CO2 per capita", x = "Energy Consumption", title = "Relationship between CO2 per capita and Energy consumption") +
  theme(plot.title = element_text(size = 7.5))
```



```{r, echo=FALSE, warning=FALSE}
(co2_pop + co2_migrant)/(co2_gdp + co2_ene)
```



These graphs display the relationship between CO2 per capita and all predictors, revealing that none of the relationships exhibit a linear pattern. As a result, linear regression is not appropriate for this dataset. Consequently, the STIRPAT framework is not suitable for this report since it is based on linear relationship between environment degradation and all predictors. Instead using non-parametric regression would be more relevant for analysing the relationship within this dataset because this regression can capture the non-linear associations between CO2 per capita and the predictors more effectively.


```{r}
# non-parametric regression without the interactions
fit_1 <- mgcv::gam(co2_capita ~ s(total_pop, bs = 'cr', k = 3) + s(net_migrant, by = total_pop, bs = 'cr', k = 3) + s(gdp_capita, by = total_pop, bs = 'cr', k = 3) + s(energy_consumption, by = total_pop, bs = 'cr', k = 3), data = co2_trans)
```



```{r}
# non-parametric regression with the interactions
fit_2 <- mgcv::gam(co2_capita ~ s(total_pop, bs = 'cr', k = 3) + s(net_migrant, by = total_pop, bs = 'cr', k = 3) + s(gdp_capita, by = total_pop, bs = 'cr', k = 3) + s(energy_consumption, by = total_pop, bs = 'cr', k = 3) + 
                     s(net_migrant, bs = 'cr', k = 3) + s(gdp_capita, bs = 'cr', k = 3) + s(energy_consumption,  bs = 'cr', k = 3), data = co2_trans)
```



```{r}
# comparing models and find the best model to explain CO2 emission in Australia
anova(fit_1, fit_2, test = "F")
```
As the result, it is significant which means this result rejects the null hypothesis that the non-parametric regression without the interactions is the best model to explain CO2 emission in Australia. Consequently, this result accept the alternative hypothesis that the non-parametric regression with the interactions is the best model to explain CO2 emission in Australia.

```{r}
summary(fit_2)
```


```{r, echo=FALSE, warning=FALSE}
nl_pop <- visreg(fit_2, "total_pop", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "Total Population",
       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and Total population") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))

nl_migrant <- visreg(fit_2, "net_migrant", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "Net Migrants",
       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and Net Migrants") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))

nl_gdp <- visreg(fit_2, "gdp_capita", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "GDP per capita",
       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and GDP per capita") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))

nl_ene <- visreg(fit_2, "energy_consumption", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "Energy Consumption",
       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and Energy Consumption") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))
```


```{r}
(nl_pop + nl_migrant)/(nl_gdp + nl_ene)
```

```{r}
fit_3 <- mgcv::gam(co2_capita ~ s(total_pop, bs = 'cr', k = 3) + s(gdp_capita, by = total_pop, bs = 'cr', k = 3) + s(energy_consumption, by = total_pop, bs = 'cr', k = 3) + s(gdp_capita, bs = 'cr', k = 3) + s(energy_consumption,  bs = 'cr', k = 3) + net_migrant:total_pop, data = co2_trans)
```

```{r}
# comparing models and find the best model to explain CO2 emission in Australia
anova(fit_2, fit_3, test = "F")
```

```{r}
summary(fit_3)
```

```{r, echo=FALSE, warning=FALSE}
nl_pop <- visreg(fit_3, "total_pop", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "Total Population",
       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and Total population") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))


nl_gdp <- visreg(fit_3, "gdp_capita", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "GDP per capita",
#       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and GDP per capita") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))

nl_ene <- visreg(fit_3, "energy_consumption", gg = TRUE) +
  theme_bw() +
  labs(y = "CO2 per capita", x = "Energy Consumption",
#       title = "The relationship in non-parametric regression",
       subtitle = "CO2 per capita and Energy Consumption") +
  theme(plot.title = element_text(size = 10),
        plot.subtitle = element_text(size = 7.5))
```


```{r, echo=FALSE}
nl_pop + nl_gdp / nl_ene
```




Technology is positive -> doesn't enough hechnology level and Furthermore, some studies such as Talbi (2017) and Le et
al (2020) sounded energy consumption illustrate technology, moreover energy consumption
measured by energy intensity, investigated energy intensity positive effects with CO2 emission
in short run. (In this case, the meaning of short run is Macroeconomics meaning)
